Trans-Dimensional Bayesian Inference for Gravitational Lens Substructures

### 2015-09-10 Papers/ms.tex
Updated some numbers and fixed marginal likelihood notation

The estimated marginal likelihood of our model
(averaged over all parameters including $N_{\rm lens}$ and $N_{\rm src}$)
is {\revisions
$\ln\left[p(\boldsymbol{D} | \boldsymbol{\theta})\right] \approx -14141.4$, and
the information (Kullback-Leibler divergence from the prior to the posterior)
is $\mathcal{H} \approx 99.4$ nats}. The information represents the degree of
compression of the posterior distribution with respect to the prior, and can
be interpreted quite literally as how much was learned about the parameters from
the data. It is also straightforward to estimate from Nested Sampling
\citep{skilling}. Its definition is:
\begin{eqnarray}
\mathcal{H} = \int p(\boldsymbol{\theta} | \boldsymbol{D})
\ln\left[\frac{p(\boldsymbol{\theta} | \boldsymbol{D})}{p(\boldsymbol{\theta})}\right]
\, d\boldsymbol{\theta}
\end{eqnarray}
and we can build intuition about its meaning based on some simple examples. One
such example is a uniform prior over a volume $V_0$ and a posterior which is
uniform over a smaller volume $V_1$ contained within $V_0$. In this case
$\mathcal{H} = \ln(V_0/V_1)$ nats. Therefore, a value of $\mathcal{H}=100$ nats
implies the posterior distribution occupies roughly $e^{-100}$ of the prior
volume.

@@

{\revisions
The estimated marginal likelihood of our model, for the `harder' simulated data,
is
$\ln\left[p(\boldsymbol{D} | \boldsymbol{\theta})\right] \approx -5671.9$, and
the information (Kullback-Leibler divergence from the prior to the posterior)
is $\mathcal{H} \approx 155.7$ nats. 
As with the easy dataset, the standardised residuals of the sampled models
resembled noise.
}

@@

The joint posterior distribution for the SIE mass total substructure mass
for the Cosmic Horseshoe is given in Figure~\ref{fig:masses3}. As with
the simulated data (Figure~\ref{fig:masses1}), we see an expected negative
correlation between these two quantities. However, the points are not as
smoothly distributed in this case (we discuss this issue further
in Section~\ref{sec:convergence}). An upper ``limit'' on the SIE mass, with
95\% posterior probability, is $5.21 \times 10^{12}$ solar masses.
The algorithm has also found
some possibilities where the substructure mass is much greater than this. In
these cases, the substructures are far from the image --- it is the environment
that is being modelled.

The posterior distribution for
$N_{\rm lens}$ (Figure~\ref{fig:N_lens3})
also shows some evidence for more than zero substructures.
In particular, the prior probability for $N_{\rm lens} > 0$ was 10/11
$\approx$ 0.91, whereas the posterior probability is approximately 0.97.
This corresponds to a Bayes Factor of approximately 3 in favour of
$N_{\rm lens} > 0$ versus the alternative $N_{\rm lens} = 0$.

@@

\begin{figure}
\begin{center}
\includegraphics[scale=0.4]{N_lens3.pdf}
\caption{{\revisions Same as Figure~\ref{fig:N_lens1}, but for the Cosmic Horseshoe.} The posterior distribution for the number of lens substructures
in the Cosmic Horseshoe system. There is mild evidence in favour of the
hypothesis that $N_{\rm lens} \neq 0$.\label{fig:N_lens3}}
\end{center}
\end{figure}

@@

The estimated marginal likelihood of our model
is {\revisions
$\ln\left[p(\boldsymbol{D} | \boldsymbol{\theta})\right] \approx -6948.5$}, and
the information (Kullback-Leibler divergence from the prior to the posterior)
is {\revisions $\mathcal{H} \approx 123.1$} nats. 
Three example models (lens and source) sampled from the posterior are shown
in Figure~\ref{fig:sources}. {\revisions
The model was able to fit the data down to the noise level.
}

@@

As a proof of concept, we demonstrated the successful recovery of a single
substructure from simulated data for which all of the model assumptions were
true. We then applied the method to an image of the Cosmic Horseshoe system
and found mild evidence in favour of the existence of a substructure, or at
the very least, a departure from a simple SIE+$\gamma$ lens profile. The
main output of our method is a set of posterior samples, each representing
a plausible scenario for the lens and the source given the data
and the assumed prior information. These samples can be displayed as a movie,
which is a very useful method for intuitively understanding the remaining
uncertainty. Any posterior summaries of interest can be
produced from the samples, but the interpretation of many of these summaries
is not necessarily straightforward. We have suggested that the
posterior expectation of the empirical measure of substructure positions is
one of the most helpful summaries.

>>>>

The estimated marginal likelihood of our model
(averaged over all parameters including $N_{\rm lens}$ and $N_{\rm src}$)
is {\revisions
$\ln\left[p(\boldsymbol{D})\right] \approx -14141.4$, and
the information (Kullback-Leibler divergence from the prior to the posterior)
is $\mathcal{H} \approx 99.4$ nats}. The information represents the degree of
compression of the posterior distribution with respect to the prior, and can
be interpreted quite literally as how much was learned about the parameters from
the data. It is also straightforward to estimate from Nested Sampling
\citep{skilling}. Its definition is:
\begin{eqnarray}
\mathcal{H} = \int p(\boldsymbol{\theta} | \boldsymbol{D})
\ln\left[\frac{p(\boldsymbol{\theta} | \boldsymbol{D})}{p(\boldsymbol{\theta})}\right]
\, d\boldsymbol{\theta}
\end{eqnarray}
and we can build intuition about its meaning based on some simple examples. One
such example is a uniform prior over a volume $V_0$ and a posterior which is
uniform over a smaller volume $V_1$ contained within $V_0$. In this case
$\mathcal{H} = \ln(V_0/V_1)$ nats. Therefore, a value of $\mathcal{H}=100$ nats
implies the posterior distribution occupies roughly $e^{-100}$ of the prior
volume.

@@

{\revisions
The estimated marginal likelihood of our model, for the `harder' simulated data,
is
$\ln\left[p(\boldsymbol{D})\right] \approx -5671.9$, and
the information (Kullback-Leibler divergence from the prior to the posterior)
is $\mathcal{H} \approx 155.7$ nats. 
As with the easy dataset, the standardised residuals of the sampled models
resembled noise.
}

@@

The joint posterior distribution for the SIE mass total substructure mass
for the Cosmic Horseshoe is given in Figure~\ref{fig:masses3}. As with
the simulated data (Figure~\ref{fig:masses1}), we see an expected negative
correlation between these two quantities. However, the points are not as
smoothly distributed in this case (we discuss this issue further
in Section~\ref{sec:convergence}). An upper ``limit'' on the SIE mass, with
95\% posterior probability, is $5.20 \times 10^{12}$ solar masses.
The algorithm has also found
some possibilities where the substructure mass is much greater than this. In
these cases, the substructures are far from the image --- it is the environment
that is being modelled.

The posterior distribution for
$N_{\rm lens}$ (Figure~\ref{fig:N_lens3})
also shows some evidence for more than zero substructures.
In particular, the prior probability for $N_{\rm lens} > 0$ was 10/11
$\approx$ 0.91, whereas the posterior probability is approximately {\revisions 0.997}.
This corresponds to a Bayes Factor of {\revisions about 300} in favour of
$N_{\rm lens} > 0$ versus the alternative $N_{\rm lens} = 0$.

@@

\begin{figure}
\begin{center}
\includegraphics[scale=0.4]{N_lens3.pdf}
\caption{{\revisions Same as Figure~\ref{fig:N_lens1}, but for the Cosmic Horseshoe.} The posterior distribution for the number of lens substructures
in the Cosmic Horseshoe system. There is {\revisions moderate} evidence in favour of the hypothesis that $N_{\rm lens} \neq 0$.\label{fig:N_lens3}}
\end{center}
\end{figure}

@@

The estimated marginal likelihood of our model
is {\revisions
$\ln\left[p(\boldsymbol{D})\right] \approx -6948.5$}, and
the information (Kullback-Leibler divergence from the prior to the posterior)
is {\revisions $\mathcal{H} \approx 123.1$} nats. 
Three example models (lens and source) sampled from the posterior are shown
in Figure~\ref{fig:sources}. {\revisions
The model was able to fit the data down to the noise level.
}

@@

As a proof of concept, we demonstrated the successful recovery of a single
substructure from simulated data for which all of the model assumptions were
true. We then applied the method to an image of the Cosmic Horseshoe system
and found {\revisions moderate} evidence in favour of the existence of a substructure, or at
the very least, a departure from a simple SIE+$\gamma$ lens profile. The
main output of our method is a set of posterior samples, each representing
a plausible scenario for the lens and the source given the data
and the assumed prior information. These samples can be displayed as a movie,
which is a very useful method for intuitively understanding the remaining
uncertainty. Any posterior summaries of interest can be
produced from the samples, but the interpretation of many of these summaries
is not necessarily straightforward. We have suggested that the
posterior expectation of the empirical measure of substructure positions is
one of the most helpful summaries.
